---
title: "p8105_hw2_sc4935"
author: "Shiwei Chen"
date: "10/2/2021"
output: github_document
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
```


# Problem 1

We first import `Trash-Wheel-Collection-Totals-8-6-19.xlsx`.

Omit non-data entries.

```{r}
TrashWheel_df = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", skip = 1, range = "A2:N535") %>% 
  janitor::clean_names() %>% 
  drop_na()

names(TrashWheel_df)
view(TrashWheel_df)

TrashWheel_df
```

Then, to omit rows that do not include dumpster-specific data, and round the number of sports balls to the nearest integer.

```{r}
TrashWheel_cleaned_df = 
  filter(TrashWheel_df, !is.na(dumpster)) %>%
  mutate(sports_balls = round(sports_balls, digits = 0))

view(TrashWheel_cleaned_df)
TrashWheel_cleaned_df
```

Read and clean precipitation data for 2018 and 2019.

```{r}
P18_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation", skip = 1, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = 2018)

P18_df
```

```{r}
P19_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2019 Precipitation", skip = 1, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = 2019)

P19_df
```

Combine precipitation datasets, convert month to a character variable.

```{r}
P1819_df = full_join(P18_df, P19_df) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year)

view(P1819_df)
P1819_df  
```

Write a paragraph about these data.

```{r}
Sportsballs_19 = filter(TrashWheel_cleaned_df, year == 2019) %>% 
  select(sports_balls)

median(pull(Sportsballs_19, sports_balls))
sum(pull(P18_df, total))

dim(TrashWheel_cleaned_df)
dim(P1819_df)

skimr::skim(TrashWheel_df)
skimr::skim(TrashWheel_cleaned_df)
skimr::skim(P1819_df)
```

In conclusion, the total precipitation in 2018 is 70.33, and the median number of sports balls in a dumpster in 2019 is 9. The Trash Wheel data set has 453 rows and 14 columns. We omit some non-data columns and rows to clean the Trash Wheel data set. For example, we round the variable of "number of sports balls" to integer. For the 2018 and 2019 Precipitation combined data set, we have 24 rows and 3 columns. I suppose that the key variable is "total", since it illustrates the monthly precipitation value for both 2018 and 2019.




# Problem 2

We first import the FiveThirtyEight data.

First, clean the data in pols-month.csv. 

```{r}
pols_month_df = read_csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(month = month.name[as.numeric(month)]) %>% 
  rename(gop = prez_gop, dem = prez_dem) %>% 
  pivot_longer(
    c(gop, dem),
    names_to = "president", 
    values_to = "win") %>%
  filter(!win == 0) %>% 
  select(-day, -win) %>% 
  view()

pols_month_df
```

Second, clean the data in snp.csv using a similar process to the above.

```{r}
snp_df = read_csv("./data/snp.csv") %>% 
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  relocate(year)

snp20_df = filter(snp_df, year <= 15) %>% 
  mutate(year = as.numeric(year) + 2000)
snp19_df = filter(snp_df, year >= 16) %>% 
  mutate(year = as.numeric(year) + 1900)

snp_df_final = bind_rows(snp20_df, snp19_df) %>% 
  mutate(year = as.character(year)) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  select(-day) %>% 
  view()
  
snp_df_final
```

Third, tidy the unemployment data so that it can be merged with the previous datasets.

```{r}
unemp_df = read_csv("./data/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month", 
    values_to = "unemployment_rate") %>% 
  janitor::clean_names() %>% 
  mutate(month = month.name[match(month,month.abb)]) %>% 
  mutate(year = as.character(year)) %>% 
  view()
  
unemp_df  
```

Merge above datasets.

```{r}
merged_df = left_join(pols_month_df, snp_df_final)
merged_df_final = left_join(merged_df, unemp_df) %>% 
  view()

merged_df_final
```

Write a short paragraph about these datasets.

```{r}
dim(merged_df_final)
range(merged_df_final$year)
names(merged_df_final)
```

In conclusion, the dataset of pols-month contains 9 variables, and illustrate the number of national politicians who are democratic or republican. In this datadet, we combine two variables, prez_gop and prez_dem, which shows whether the president was republican on the associated date clearly. For dataset of snp, it contains 2 variables, date and the closing values of the S&P stock index on the associated date. For unemployment dataset, it presents the monthly unemployment rate start on 1948. 

For the resulting dataset, the dimention is 822 and 11, which implies 822 rows and 11 columns, the range of year from 1947 to 2015, and there are some key variables, like year, month, president, close, and unemployment_rate. From original dataset, the date variable is seperated to year and month. President shows whether he is democratic or republican. Close and unemployment_rate are 2 most important value to reflect the society.





# Problem 3

We first import the data of the popularity of baby names.

Load and tidy the data, and remove duplicate rows.

```{r}
popular_baby_names_df = read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "WHITE NON HISP" = "WHITE NON HISPANIC", "BLACK NON HISP" = "BLACK NON HISPANIC")) %>% 
  mutate(gender = str_to_sentence(gender),
         ethnicity = str_to_sentence(ethnicity),
         childs_first_name = str_to_sentence(childs_first_name)) %>% 
  mutate(year_of_birth = as.character(year_of_birth)) %>% 
  distinct() %>% 
  view()

popular_baby_names_df
```

Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time, and produce a similar table showing the most popular name among male children over time.

```{r}
Olivia_as_female_name = filter(popular_baby_names_df, childs_first_name == "Olivia", gender == "Female") %>% 
  select( -gender, -childs_first_name, -count) %>% 
  pivot_wider(
  names_from = "year_of_birth", 
  values_from = "rank") %>% 
  view()

Olivia_as_female_name


Most_popular_male_name = filter(popular_baby_names_df, rank == 1, gender == "Male") %>%
  select(-gender, -count, -rank) %>% 
  pivot_wider(
  names_from = "year_of_birth", 
  values_from = "childs_first_name") %>% 
  view()

Most_popular_male_name
```

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}
Male_white_2016 = filter(popular_baby_names_df, gender == "Male", year_of_birth == 2016, ethnicity == "White non hispanic")

Male_white_2016

ggplot(Male_white_2016, aes(x = rank, y = count)) + geom_point() 
  
  
```

Finished homework.





